On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit

### 2023-04-30 manuscript/manuscript.tex
minor

\jy{how about removing the panels for $-0.1$ and 0.1?}
\eds{seems like too few if those are removed. How about $\{-0.4,-0.2,0.2,0.4\}$
or $\{-0.30,-0.15,0.15,0.30\}$? Is it easy/quick to generate the plots?}

@@

The invalidity of the standard statistic for serially dependent data can be
illustrated by a simple simulation study. Consider a first-order autoregressive
or AR(1) model with marginal normal (or gamma) distribution. For each AR
coefficient $\psi \in \{-0.3, -0.1, 0.1, 0.3\}$, we generated 1000
series of length 200. For each series, we applied the KS statistic
in~\eqref{eq:ks_standard} to test $H_0$ that the $X_i$'s follow $N(8, 8)$ (or 
$\Gamma(8,1)$) marginally. The PP-plots of the p-values from $1000$ replicates 
for all the $\psi$ values are displayed in Figure~\ref{fig:pp_s}. As 
$\phi$ diverges from 0, the p-values look less likely to be generated from a 
standard uniform distributions. Positive $\psi$ values led to p-values with a 
distribution function higher than the standard uniform distribution function as 
there were more smaller p-values than expected, and 
hence, %conservative 
liberal tests. Negative $\psi$ values led to p-values with a 
distribution function lower than the standard uniform distribution as there 
were more larger p-values than expected, and hence, 
%liberal 
conservative tests. 
\eds{the previous two sentences are not consistent with what is 
written at the beginning of this section. If I am interpreting the plots 
correctly, then positive $\psi$ leads to p-values that are too low which would
correspond to liberal tests (consistent with what was written at the beginning 
of this section).} \jy{Need to rephrase. A distribution function higher than
  another distribution means that it has more smaller values.}
\eds{do the edits fix the problem?}

@@

\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the working ARMA$(p, q)$ model with
  innovation variance $1 / \sigma^2$ such that the $Z_i$'s are marginally
  standard normal variables.
\item
  Form a bootstrap sample
  $X_i^* = F_{\hat\theta_n}^{-1} [\Phi(Z_i^*)]$, $i = 1, \ldots, n$, where
  $F_{\hat\theta_n}^{-1}$ is the quantile function of $F_{\hat\theta_n}$.
\item
  Fit $F_\theta$ to $X_1^*, \ldots, X_n^*$ to obtain estimate $\hat\theta_n^*$
  of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x \lvert F_n^* (x) - F_{\hat\theta_n}^* (x) \rvert.
  \]
	\eds{check placement of * in second F}
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of the $B$ test statistics to approximate
  the null distribution of the observed statistic.
\end{enumerate}
The p-value of $D_n$ is approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Compared to the algorithm in
the last section, here we use $F_{\hat\theta_n}$ and $F_{\hat\theta_n}^*$ in
place of the known~$F$. \eds{check placement of * in second F}

@@

Figure~\ref{fig:pp_ssf} shows the PP-plots resulting p-valuess of the
procedure performed on data generated from same simulation settings as in
Section~\ref{sec:dependence}. For comparison, we also displayed the PP-plot of
a naive bootstrap procedure that adjusts for the unspecified parameters as in
Section~\ref{sec:fitted} but not for the serial dependence. 
The super naive approach that does not adjust for fitted parameters nor serial
dependence would have double sin, so the results are not reported. 
The p-values from
naively fitting parameters but not adjusting for dependence clearly deviates
from the standard uniform distribution except in the MA(1) case.
\jy{needs an explanation ...}
\eds{if we had used a different MA parameter other than 0.8, would we see  
similar behavior?} 
Similar to Section~\ref{sec:dependence}, the working normal copula approach
gives p-values with the desired standard uniform distribution regardless of the
true dependence.

@@

Finally, we report a small study of the power of the one-sample KS test with
unspecified parameters and serially dependent data. Using the three dependence
models characterized by ARMA(1, 1), AR(2), and MA(1), we generated serially
dependent data with marginal distributions $N(8, 8)$ left truncated by zero and
$\Gamma(8, 1)$. 
\eds{same AR and MA parameters as the previous section?}
Then, for data from $\Gamma(8, 1)$, we tested $H_0$ that the
data has marginally a normal distribution; for data from $N(8, 8)$ left
truncated by zero, we tested $H_0$ that the data has marginally a gamma
distribution;  In both cases, no parameters were specified. Two sample sizes
were considered, $n \in \{100, 200\}$. Table~\ref{tab:power} summarizes the
power of rejecting the null hypothesis at significance level~0.05. The tests
based on a working copula have substantial power in rejecting the null
hypothesis and the power increases as sample size increases. The power from the
naive tests where the serial dependence is ignored, although the unspecified
parameters are accounted for (i.e., procedure from 
Section~\ref{sec:dependence}), appear to 
have higher power than the proposed tests, but this is not reliable since they 
do not hold their sizes as demonstrated in the Figures.

@@

The KS test has base assumptions that the hypothesized distribution is
completely specified and the data is independent. When these assumptions are
violated, the test is no longer accurate and remedies must be performed. In the
case of fitted parameters, parametric and non-parametric bootstrap can restore
the size of the test. A bias correction is required if the non-parametric form
is used \citep{babu2004goodness}. In the case of serially dependent data, 
a procedure using bootstrap with copulas to model dependency shows positive
results. When both assumptions are violated, i.e., where the data has serial 
dependence and parameters must be fitted, an adjusted copula procedure also 
shows positive results. The tests appear effective for a variety of families of 
distributions. The copula remedy is not a complete solution and has limitations. 
Regardless of the true dependence, we assume an ARMA($p,q$) dependence structure. 
Therefore, if the assumed dependence structure is a close approximation of the
truth, the approach can work. However, if the approximation is too far from the 
true dependence structure the approach does not completely remedy the issue. 
As well, tests were only performed with the normal copula. It is possible that 
other copulas could provide stronger results based on the true dependence of the 
data.

>>>>

\jy{how about removing the panels for $-0.1$ and 0.1?}
\eds{seems like too few if those are removed. How about $\{-0.4,-0.2,0.2,0.4\}$
or $\{-0.30,-0.15,0.15,0.30\}$? Is it easy/quick to generate the plots?} 
\jy{Will get jobs done for $\{-0.4,-0.2,0.2,0.4\}$}

@@

The invalidity of the standard statistic for serially dependent data can be
illustrated by a simple simulation study. Consider a first-order autoregressive
or AR(1) model with marginal normal (or gamma) distribution. For each AR
coefficient $\psi \in \{-0.3, -0.1, 0.1, 0.3\}$, we generated 1000
series of length 200. For each series, we applied the KS statistic
in~\eqref{eq:ks_standard} to test $H_0$ that the $X_i$'s follow $N(8, 8)$ (or 
$\Gamma(8,1)$) marginally. The PP-plots of the p-values from $1000$ replicates 
for all the $\psi$ values are displayed in Figure~\ref{fig:pp_s}. As 
$\phi$ diverges from 0, the p-values look less likely to be generated from a 
standard uniform distributions. Positive $\psi$ values led to p-values with a 
distribution function higher than the standard uniform distribution function as 
there were more smaller p-values than expected, and 
hence, %conservative 
liberal tests. Negative $\psi$ values led to p-values with a 
distribution function lower than the standard uniform distribution as there 
were more larger p-values than expected, and hence, 
%liberal 
conservative tests. 
\eds{the previous two sentences are not consistent with what is 
written at the beginning of this section. If I am interpreting the plots 
correctly, then positive $\psi$ leads to p-values that are too low which would
correspond to liberal tests (consistent with what was written at the beginning 
of this section).} \jy{Need to rephrase. A distribution function higher than
  another distribution means that it has more smaller values.}
\eds{do the edits fix the problem?}
\jy{can we rephrase ``than excepted'' to ``than there should be'' or something
  along the line?}

@@

\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the working ARMA$(p, q)$ model with
  innovation variance $1 / \sigma^2$ such that the $Z_i$'s are marginally
  standard normal variables.
\item
  Form a bootstrap sample
  $X_i^* = F_{\hat\theta_n}^{-1} [\Phi(Z_i^*)]$, $i = 1, \ldots, n$, where
  $F_{\hat\theta_n}^{-1}$ is the quantile function of $F_{\hat\theta_n}$.
\item
  Fit $F_\theta$ to $X_1^*, \ldots, X_n^*$ to obtain estimate $\hat\theta_n^*$
  of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x \lvert F_n^* (x) - F_{\hat\theta_n}^* (x) \rvert.
  \]
  \eds{check placement of * in second F}
  \jy{the start should be with hat theta from step 3}
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of the $B$ test statistics to approximate
  the null distribution of the observed statistic.
\end{enumerate}
The p-value of $D_n$ is approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Compared to the algorithm in
the last section, here we use $F_{\hat\theta_n}$ and $F_{\hat\theta_n}^*$ in
place of the known~$F$. \eds{check placement of * in second F}

@@

Figure~\ref{fig:pp_ssf} shows the PP-plots resulting p-valuess of the
procedure performed on data generated from same simulation settings as in
Section~\ref{sec:dependence}. For comparison, we also displayed the PP-plot of
a naive bootstrap procedure that adjusts for the unspecified parameters as in
Section~\ref{sec:fitted} but not for the serial dependence. 
The super naive approach that does not adjust for fitted parameters nor serial
dependence would have double sin, so the results are not reported. 
The p-values from
naively fitting parameters but not adjusting for dependence clearly deviates
from the standard uniform distribution except in the MA(1) case.
\jy{needs an explanation ...}
\eds{if we had used a different MA parameter other than 0.8, would we see  
similar behavior?}  \jy{yes. I have more results for other MA coefficients to
  show when we meet.}
Similar to Section~\ref{sec:dependence}, the working normal copula approach
gives p-values with the desired standard uniform distribution regardless of the
true dependence.

@@

Finally, we report a small study of the power of the one-sample KS test with
unspecified parameters and serially dependent data. Using the three dependence
models characterized by ARMA(1, 1), AR(2), and MA(1) in the simulation settings
in the last section, we generated serially
dependent data with marginal distributions $N(8, 8)$ left truncated by zero and
$\Gamma(8, 1)$. 
\eds{same AR and MA parameters as the previous section?} \jy{yes; edited.}
Then, for data from $\Gamma(8, 1)$, we tested $H_0$ that the
data has marginally a normal distribution; for data from $N(8, 8)$ left
truncated by zero, we tested $H_0$ that the data has marginally a gamma
distribution;  In both cases, no parameters were specified. Two sample sizes
were considered, $n \in \{100, 200\}$. Table~\ref{tab:power} summarizes the
power of rejecting the null hypothesis at significance level~0.05. The tests
based on a working copula have substantial power in rejecting the null
hypothesis and the power increases as sample size increases. The power from the
naive tests where the serial dependence is ignored, although the unspecified
parameters are accounted for (i.e., procedure from 
Section~\ref{sec:dependence}), appear to 
have higher power than the proposed tests, but this is not reliable since they 
do not hold their sizes as demonstrated in the Figures.

@@

The KS test has base assumptions that the hypothesized distribution is
completely specified and the data is independent. When these assumptions are
violated, the test is no longer accurate and remedies must be performed. In the
case of fitted parameters, parametric and non-parametric bootstrap can restore
the size of the test. A bias correction is required if the non-parametric form
is used \citep{babu2004goodness}. In the case of serially dependent data, 
a procedure using bootstrap with the normal copula determined by an
automatically selected ARMA model to account for dependency shows positive
results. When both assumptions are violated, i.e., where the data has serial 
dependence and parameters must be fitted, an adjusted copula procedure also 
shows positive results. The tests appear effective for a variety of families of 
distributions. The copula remedy is not a complete solution and has limitations. 
Regardless of the true dependence, we assume an ARMA($p,q$) dependence structure. 
Therefore, if the assumed dependence structure is a close approximation of the
truth, the approach can work. However, if the approximation is too far from the 
true dependence structure the approach does not completely remedy the issue. 
As well, tests were only performed with the normal copula. It is possible that 
other copulas could provide stronger results based on the true dependence of the 
data.
